Unbalanced Mallows Models for Optimizing Expensive Black-Box Permutation Problems

### 2021-02-04 paper/main.tex
plots
minor

In Section~\ref{sec:permutations}, we highlighted the important difference between rankings and orderings, usually neglected in the optimization literature. Statistical models (Borda or the \ken distance) are usually defined for rankings while the optimization literature usually considers orderings. For example, the Swap distance $d_s(\sigma,\pi)$ is the \ken distance counterpart in optimization, $d_s(\sigma,\pi) = d(\sigma^{-1}, \pi^{-1})$. The concepts of rankings and orderings are usually confused of ignored. An example of such confusion is the claim that Swap distance is the same metric as  the \ken distance, which is false except for particular permutations. In reality, the \emph{swap} distance is the \ken distance of the inverted permutations.

As a consequence of this important distinction, when evaluating classical
combinatorial problems, we must convert rankings to orderings by inverting the
permutation, as done in Eq.~\ref{eq:lop} and Eq.~\ref{eq:pfsp}, for the LOP and
PFSP, respectively. This might seem an implementation detail but a failure to use the correct interpretation has grave consequences.

@@

\begin{figure*}[tb]
  \small
  \centering%
    \includegraphics[width=0.49\textwidth]{../img/fitness_lop_IO_N-t59b11xx}
    \includegraphics[width=0.49\textwidth]{../img/fitness_lop_SGB_N-sgb75_02}
\\
    \includegraphics[width=0.49\textwidth]{../img/fitness_lop_RandB_N-p40-01}
    \begin{minipage}{0.49\linewidth}
          \texttt{RandB N-p50-01} (LOP)\\[-0.5ex]
    \includegraphics[width=\textwidth]{../img/fitness_lop_RandB_N-p50-01}
  \end{minipage}
    \\
    \begin{minipage}{0.49\linewidth}
    \texttt{reC19} (PFSP)\\[-0.5ex]
\includegraphics[width=\textwidth]{../img/fitness_pfsp_rec19_txt}
\end{minipage}
\begin{minipage}{0.49\linewidth}
    \texttt{reC31} (PFSP)\\[-0.5ex]
    \includegraphics[width=\textwidth]{../img/fitness_pfsp_rec31_txt}
  \end{minipage}
\label{fig:results}
  \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on various instances.\label{fig:rec05}}
\end{figure*}

>>>>

In Section~\ref{sec:permutations}, we highlighted the important difference between rankings and orderings, usually neglected in the optimization literature. Statistical models (Borda or the \ken distance) are usually defined for rankings while the optimization literature usually considers orderings. For example, the Swap distance $d_s(\sigma,\pi)$ is the counterpart of the  \ken distance  in optimization, $d_s(\sigma,\pi) = d(\sigma^{-1}, \pi^{-1})$. The concepts of rankings and orderings are usually confused of ignored. An example of such confusion is the claim that Swap distance is the same metric as  the \ken distance, which is false except for particular permutations. In reality, the \emph{swap} distance is the \ken distance of the inverted permutations.

As a consequence of this important distinction, when evaluating classical
combinatorial problems, UMM transforms rankings to orderings by inverting the
permutation, as done in Eq.~\ref{eq:lop} and Eq.~\ref{eq:pfsp}, for the LOP and
PFSP, respectively. This might seem an implementation detail but a failure to use the correct interpretation has grave consequences.

@@

\begin{figure*}[tb]
\small
\centering%
\begin{minipage}{0.49\linewidth}
  \texttt{N-t59b11xx} (LOP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/fitness_lop_IO_N-t59b11xx}
\end{minipage}
\begin{minipage}{0.49\linewidth}
  \texttt{N-sgb75-01} (LOP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/fitness_lop_SGB_N-sgb75_02}
\end{minipage}
\\
\begin{minipage}{0.49\linewidth}
  \texttt{N-p40-01} (LOP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/fitness_lop_RandB_N-p40-01}
\end{minipage}
\begin{minipage}{0.49\linewidth}
  \texttt{N-p50-01} (LOP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/fitness_lop_RandB_N-p50-01}
\end{minipage}
\\
\begin{minipage}{0.49\linewidth}
  \texttt{reC19} (PFSP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/fitness_pfsp_rec19_txt}
\end{minipage}
\begin{minipage}{0.49\linewidth}
  \texttt{reC31} (PFSP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/fitness_pfsp_rec31_txt}
\end{minipage}
\label{fig:results}
\caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on various instances.\label{fig:rec05}}
\end{figure*}
